---
title: "MA5832 Capstone Assessment (Assessment 3) RMarkdown all"
author: "Samuel Bousfield"
date: "`r Sys.Date()`"
output:
  word_document: 
    fig_caption: yes
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.width = 10)
```

```{r Working Directory, include=FALSE}
setwd("C:/Users/sbous/OneDrive/ASUS Zenbook Pro Duo Folders/Master of Data Science/Data Mining and Machine Learning/Assessment 3")
```

```{r libraries}
library(tidyverse)
library(imputeTS)
library(caret)
library(neuralnet)
library(e1071)
```

```{r Data, include=FALSE}
AUS_Data <- readxl::read_excel("AUS_Data.xlsx", skip = 1)
#head(AUS_Data)
colnames(AUS_Data)
colnames(AUS_Data) <- c("Period", "Unemployment Rate (%)", "GDP (% change)", "Government Final Expenditure (% change)", "All Sector Final Expenditure (% change)", "Term of Trade Index (%)", "All Group CPI", "Job Vacancies (x1000)", "Estimated Resident Population (x1000)")
colnames(AUS_Data)
```

# Overview of Unemployment Rate

Unemployment is a critical problem within a society, leading to poverty, low living standards, pressure on welfare and other public services, and is proven to lead to worse health outcomes, physically and mentally. The goal of this report is to assess over 20 years of unemployment data, and seven socio-economic predictors, and apply a Support Vector Machine (SVM) model and an artificial neural network (ANN), to assess the suitability of using these predictors to estimate unemployment rate. The graph below shows the correlation between each predictor variable, and the Unemployment Rate (%). This graph shows that the Job Vacancies predictor is the best predictor of the unemployment rate, followed by Estimated Resident Population and All Group CPI. Government Final Expenditure appears to have almost no correlation whatsoever, this variable also has a positive correlation. The initial implications of this are that changing government expenditure does not really have an effect on the unemployment rate, although an increase in government expenditure is correlated with an extremely small increase in the unemployment rate.

```{r Unemployment Rate 1999-2020}
unemployment <- AUS_Data[72:158,2:ncol(AUS_Data)]
for (i in 1:ncol(unemployment)){
 unemployment[[i]] <- as.numeric(unemployment[[i]])
}
unemployment_cor <- as.data.frame(cor(unemployment, use = "pairwise.complete.obs"))
unemployment_cor <- unemployment_cor[1,]
unemployment_cor <- pivot_longer(unemployment_cor, cols = -`Unemployment Rate (%)`, names_to = "variable", values_to = "correlation")

ggplot(unemployment_cor) +
 aes(x = variable, weight = correlation) +
 geom_bar(fill = "lightgrey") +
 labs(x = "Predictor Variables", 
 y = "Correlation", title = "Figure 1. Predictor Correlation to Unemployment Rate (%)") +
 geom_text(aes(label = sprintf("%.4f", correlation), y = correlation, hjust = ifelse(variable == "Government Final Expenditure (% change)",1, -0.1)))+
 coord_flip() +
 theme_minimal()
```

# Data

## 2(a.1) - Preprocessing
Preprocessing included assessing the column types. A brief examination of the data showed that all the data is numeric, assessing the column types showed that two of the columns are of type character. This is caused by missing data. The predictor and response variables were then coerced to be of type numeric, which introduced NAs. there are 5 missing values in the Job Vacancies and Estimated Resident Population fields, which is approximately 3.2% of the data in each column, although these are at different row levels, therefore approximately 6.3% of the data has missing values. Three main options are available.
1) Since the observations with missing data accounts for less than 10% of the data, they could be dropped from the data set. This is not a good idea as the testing datasets for the models have been specifically defined, and this would drop ~50% of the testing data
2) The missing values could be left as is.
3) Interpolate or Impute values.
These three options will be explored in the next.
The summary of the data shows that the data covers the period from June 1981 to September 2020. The unemployment rate for this period is between 4.1% and 11.133%, with a mean employment rate of 6.852%. GDP change ranges includes negative and positive values, ranging from -7% change to 3.3% change.  Government and All Sector Expenditure % change also covers negative and positive values with a minimum of -4.6% and -8.3% change and a maximum of 7.5% and 5.9% change respectively, likewise the Term of Trade Index % covers a range of -8.1 to 5.9%. These values indicate there is likely to be interactions within the data between these predictors and the response variable.

```{r Minor Preprocessing}
#, eval=FALSE, include=FALSE}
for (i in 1:ncol(AUS_Data)){
  print(typeof(AUS_Data[[i]]))
}
#All variables should be numeric (except for date), must convert from character
#create new data set to avoid messing with old dataset
data <- AUS_Data
for (i in 2:ncol(data)){
 data[[i]] <- as.numeric(data[[i]])
}
#warning that NA's were introduced in conversion process in previous loop.
#check to see count of and % of NA's in each column
colSums(is.na(data))
colSums(is.na(data))/nrow(data) * 100
#5 errors present in 2 of the columns (~3.16%). check to see which rows these are in
which(is.na(data), arr.ind=TRUE)
(10 / nrow(data)) * 100
#10 total rows are missing data, this is approximately 6.3% of the data
summary(data)
```

```{r Transform Data into Numeric}
data <- AUS_Data
for (i in 2:ncol(data)){
 data[[i]] <- as.numeric(data[[i]])
}
```

## 2(a.2) - Missing Value Imputation
There are two variables with missing values: Estimated Resident
Population (x1000), and Job Vacancies (x1000). Examining a basic plot of
the two variables, the population variable appears to be increasing
relatively steady. One method of missing value imputation is to find an
external source of data, which should be easy in this case, population
statistics are usually pretty easy to find, however, this is
inappropriate in two ways for this data, 1) this dataset is treating the
population of Australia as x10 greater than the real variables 2) even
though this is fairly obviously the case, because this is not the real
data, it cannot be properly inferred that we can take the real
population statistics and multiply by 10. This is also the case for the
Job Vacancies variable.

```{r Plot Variables with missing values}
plot(data$`Estimated Resident Population (x1000)`, main = "Figure 2. Estimated Resident Population (x1000) real values")
plot(data$`Job Vacancies (x1000)`, main = "Figure 3. Job Vacancies (x1000) real values")
```

Examining the plots above for the variables with missing values, we can
see that the Estimate Resident Population appears to be increasing in a
linear fashion, and while the Job Vacancies variable is not linear, it
also appears to be following a pattern, likely a cubic function. Thus,
some of the normal methods for dealing with missing data are likely to
be inappropriate (knn, mean imputation). The best static value
imputation for these is likely to be the previous filled value (or max
value) in the case of the resident population, and either that or the
mean value of the previous filled value and following filled value for
the Job Vacancies data. However the most appropriate option for both is
likely to interpolate. A few methods will be assessed for appropriateness, using the na_interpolation() function from imputeTS package, and in generating a simple linear or polynomial model. These will then be examined for what appears to fit the data best, and then the values imputed into the original data.

```{r Test imputation methods}
data_1 <- data[,c("Job Vacancies (x1000)", "Estimated Resident Population (x1000)")]
data_1$index <- 1:nrow(data_1)
data_1$job_stineimpute <- na_interpolation(data_1$`Job Vacancies (x1000)`, option = "stine")
data_1$job_linearimpute <- na_interpolation(data_1$`Job Vacancies (x1000)`, option = "linear")
data_1$job_splineimpute <- na_interpolation(data_1$`Job Vacancies (x1000)`, option = "spline")

data_1$erp_stineimpute <- na_interpolation(data_1$`Estimated Resident Population (x1000)`, option = "stine")
data_1$erp_linearimpute <- na_interpolation(data_1$`Estimated Resident Population (x1000)`, option = "linear")
data_1$erp_splineimpute <- na_interpolation(data_1$`Estimated Resident Population (x1000)`, option = "spline")

linear_model <- lm(data$`Estimated Resident Population (x1000)`~ index, data=data_1)
predicted_values_linear <- predict(linear_model, newdata = data_1)
data_1$erp_linearmodel <- data_1$`Estimated Resident Population (x1000)`
data_1$erp_linearmodel[is.na(data_1$erp_linearmodel)] <- predicted_values_linear[is.na(data_1$erp_linearmodel)]

poly_model <- lm(data$`Estimated Resident Population (x1000)`~ poly(index, 2), data=data_1)
predicted_values <- predict(poly_model, newdata = data_1)
data_1$erp_polymodel <- data_1$`Estimated Resident Population (x1000)`
data_1$erp_polymodel[is.na(data_1$erp_polymodel)] <- predicted_values[is.na(data_1$erp_polymodel)]

#pivot longer
data_long_job <- pivot_longer(
  data_1,
  cols = c(`Job Vacancies (x1000)`, job_stineimpute, job_linearimpute, job_splineimpute),
  names_to = "Original Field",
  values_to = "Value"
    )
data_long_job <- data_long_job[, c("index", "Original Field", "Value")]

data_long_erp <- pivot_longer(
  data_1,
  cols = c(`Estimated Resident Population (x1000)`, erp_stineimpute, erp_linearimpute, erp_splineimpute, erp_linearmodel, erp_polymodel),
  names_to = "Original Field",
  values_to = "Value"
    )
data_long_erp <- data_long_erp[, c("index", "Original Field", "Value")]
```

The methods have been completed, these will be assessed visually using the ggplot package. The geom_jitter option has been used so that indiviudal points can be seen more clearly. This offset should be taken into account when assessing the best fitt for the data.

```{r Visualise Imputations}
ggplot(data_long_erp) +
 aes(x = index, y = Value, colour = `Original Field`) +
 geom_jitter(size = 1.5) +
 scale_color_hue(direction = 1) +
 labs(title = "Figure 4. Estimated Resident Population",
      subtitle = "Real and Interpolated Values") +
 theme_minimal()


ggplot(data_long_job) +
 aes(x = index, y = Value, colour = `Original Field`) +
 geom_jitter(shape = "circle", 
 size = 1.5) +
 scale_color_hue(direction = 1) +
 labs(title = "Figure 5. Job Vacancies",
      subtitle = "Real and Interpolated Values") +
 theme_minimal()
```

We can see that in the Estimated Resident Population, the slight curve
makes a significant difference, resulting in the linear model not
fitting the data well, and the other imputation methods just take the
max value in the field. However, the polynomial model fits the data very
well.

In the Job Vacancies data, the linear and Stine imputation both appear
to work almost equally well, while the spline imputation takes the
lowest value beyond the next observation with a value. While this may be
correct in the real data, it is better to assume a value between the two
values with data. In this case, the stine imputation option has been chosen, although the linear option provides a similar level of appropriateness.

```{r impute values into original data and drop testing data frames}
#impute
data$`Job Vacancies (x1000)` <- data_1$job_stineimpute
data$`Estimated Resident Population (x1000)` <- data_1$erp_polymodel
#check to ensure no more NA's, and plots look good
colSums(is.na(data))

plot(data$`Estimated Resident Population (x1000)`, main = "Figure 6. Estimated Resident Population (x1000) Values Imputed")
plot(data$`Job Vacancies (x1000)`, main = "Figure 7. Job Vacancies (x1000) Values Imputed")

#drop unnecessary things from environment
rm(data_1, data_long_erp, data_long_job, linear_model, poly_model)
```

## 2(b) - Descriptive Statistics

# Machine Learning

Because of the choice of ML algorithm, and the comparison to a neural
network, it is good practice to scale the data for a few reasons: . We
will also be removing the "Period" variable, as this is more of an
identifier, rather than

Testing data is specified to be data from March 2018 - September 2020,
therefore Training data is data up to 2017

```{r Split Training and Test Data}
#scale training data
data_scaled <- data.frame(scale(data[, 3:9]))
#rebind the response variable, and rename the columns
data_scaled <- cbind(data$Period ,data$`Unemployment Rate (%)`, data_scaled)
colnames(data_scaled) <- c("x0", "Y", "X1", "X2", "X3", "X4", "X5", "X6", "X7")

#create data partition, and split into training and testing data
training_data <- data_scaled[1:147,]
testing_data <- data_scaled[148:nrow(data),]
model_formula <- Y ~ X1 + X2 + X3 + X4 + X5 + X6 + X7
```

## 3(a) - Justify Choice

SVM has been chosen because this is a regression model, and all the variables are numeric. This means that there is likely a complex relationship between all 7 of the predictor variables, and the response variable. This is an area in which an SVM with a non-linear kernel shines over Forests. Another major reason for choosing an SVM model over a Forest is the sample size. SVMS perform well with small and medium sized datasets, while Forest's may overfit.

## 3(b) - Hyperparameters

Because of the small dataset, the Cost and Sigma values can be tuned in a reasonable amount of time. The C (cost) hyperparameter represents the trade-off between training error and testing error. A higher value here encourages the SVM to classify more training points accurately, resulting in a smaller margin, but has the potential trade off of overfitting. The Sigma hyper parameter influences the boundary. A smaller sigma is sensitive to local patterns, while a larger sigma is more sensitive to global patterns, this may lead to overfitting if it is too small, or underfitting if it is too large.

Using the cross-validation method, the hyperparameters will be tuned, the number of cross-validations and repeats can be relatively large, due to the small number of observations. The optimal hyperparameters will then be taken and used to train another model, because of the high number of cross-validations and repeats, these optimal hyperparameters can be trusted to be the optimal combination of hyperparameters in this case.

```{r SVM Hyperparameter Tuning}
#C_values and Sigma
# tuning_time_svm <- system.time({
# c_values <- c(0.1, 0.5, 1, 2.5, 5, 10)
# Sigma_values <- c(0.1, 0.5, 1, 2.5, 5, 10)
# grid <- expand.grid(sigma = Sigma_values, C = c_values)
# ctrl_method <- trainControl(method = "repeatedcv", number = 100, repeats = 10)
# set.seed(0)
# hyperparameter_tuning_svm <- train(model_formula
#                        , data = training_data
#                        , method = "svmRadial"
#                        , trControl = ctrl_method
#                        , tuneGrid = grid)
# })
# 
# saveRDS(hyperparameter_tuning_svm, file = "hyperparameter_tuning_svm.rds")
# saveRDS(tuning_time_svm, file = "tuning_time_svm.rds")
#commented out to prevent rerunning
```

The hyperparameter tuning has been saved to RDS to be reloaded, this is due to training time. The optimal hyperparameters will then be extracted from the tuning model, and used to train the final model. Which will also be saved, due to training time.

``` {r generate SVM model}
hyperparameter_tuning_svm <- readRDS(file = "hyperparameter_tuning_svm.rds")
#hyperparameter_tuning_svm
best_sigma <- hyperparameter_tuning_svm$bestTune$sigma
best_cost <- hyperparameter_tuning_svm$bestTune$C
time_svm_tuning <- readRDS(file = "tuning_time_svm.rds")
#tuning_time_svm
# set.seed(0)
# time_svm <- system.time({svm_model <- svm(model_formula,
#                                                  data = training_data,
#                                                  kernel = "radial",
#                                                  sigma = best_sigma,
#                                                  cost = best_cost)
#                        })
# saveRDS(time_svm, file = "time_svm.rds")
# saveRDS(svm_model, file = "svm_model.rds")
svm_model <- readRDS(file = "svm_model.rds")
time_svm <- readRDS(file = "time_svm.rds")
```


```{r Prediction Results Function}
#After multiple runs through, of using the SAME lines of code to get the regression results, I decided to turn it into a function #DRY
predict_funct <- function(model, estimate_results, real_results){
    x <- predict(model, newdata = estimate_results)
    y <- real_results
    postResample(x, y)
  }
```

## 3(c) - Predictive Performance on Training Data

```{r assess model on training data}
#load in best hyperparameters
# Original Lines of Code
# svm_predict_train <- predict(hyperparameter_tuning_svm, newdata = training_data)
# #assess the svm model accuracy on the training data using the postResample() function in caret
# svm_accuracy_train <- postResample(svm_predict_train, training_data$`Unemployment Rate (%)`)
# svm_accuracy_train
# Use my function
print("SVM Model on Training Data")
predict_funct(svm_model, training_data, training_data$Y)
```

These values are very good, it is difficult to evaluate the Root Mean
Square Error (RMSE) in a vacuum, unless it is used to compare to other
models, so the particular value will not be examined too deeply, but a
lower RMSE indicates a model that fits the data better. The Rsquared
value 0.92263239 is an exceptionally good value. The Rsquared value indicates
the amount of variance in the data/response variable that is explained
by the model (0 = 0%; 1 = 100%). Thus this value indicates that \~92% of
the variance in the data is explained by the model, which indicates a
model of exceptional fit. The Mean Absolute Error (MAE), like the RMSE,
is an average of the errors and a lower value indicates a better model;
however, because the MAE value is the absolute values, the result is in
the original units of the predictor variable. This means that the
average difference between what the model predicts the Unemployment Rate
(%) to be, and the actual Unemployment Rate (%) is 0.3141412%. This
could be considered a very good value, the range of real results is
between 4.1% and 11.133%, with a mean of 6.852% and a median of 6.250%.
This indicates that the SVM model varies from the average of the real
results by \~4.5%. It is important to note that these predictions are on
the training data, where the model would be expected to perform well,
although a poor result here is a good indicator of model unsuitability.
Next the SVM model will be assessed on the training data.

## 3(d) - Predictive Performance on Testing Data

```{r assess model on testing data}
print("SVM Model on Testing Data")
predict_funct(svm_model, testing_data, testing_data$Y)
```

The model has performed very well on the test data, interestingly the
RMSE and Rsquared values indicate that the model is a better fit on the
test data than the training data (Train RMSE = 0.5039870, Test RMSE =0.3822914; Train
Rsquared = 0.9226239, Test Rsquared = 0.9414006), although the MAE is larger(0.3310902) as expected.
These results are likely the function of the small testing data set.

These values indicate that while the model fits the data better, the
incorrect values are slightly more incorrect. Although this only results
in the error from the average of the real results in \~4.8%. Again it is critical to recognise that the testing data is an extremely small dataset (11 observations).

The results from both sets of predictions indicate that the svm model performs exceptionally well.


# Neural Network

## 4(a.1) - Describe the Structure of the Model
An Artificial Neural Network (ANN),  

## 4(a.2) - Implement the model
Implementation of the initial ANN will be performed using 7 hidden layers of 1 neuron each. These numbers were chosen semi-randomly. 7 layers were chosen because the data has 7 predictor variables and 1 layer was chosen because one "variable" (date) is not a predictor or response variable. It is important to note that these values are "random". This model could be trained with any number of hidden layers and any number of neurons, the only limit would be training time.
```{r Initial Neural Network Model}
#commented out to prevent rerunning
# #2 hidden layers of 7 neurons in first hidden layer, 1 neuron in second hidden layer
# initial_ann_training_time <- system.time({
#     set.seed(0)
#     initial_ann <- neuralnet(model_formula,
#                              data = training_data,
#                              hidden = c(7,1),
#                              stepmax = 1e7)
# })
# # saveRDS(initial_ann_training_time, file = "initial_ann_training_time.rds")
# # saveRDS(initial_ann, file = "initial_ann.rds")

time_ann <- readRDS(file = "initial_ann_training_time.rds")
initial_ann <- readRDS(file = "initial_ann.rds")
```

## 4(b) - Predictive Performance on Training Data

```{r test model on training data}
print("Initial Neural Network on Training Data:")
predict_funct(initial_ann, training_data, training_data$Y)
```
These results are phenomenal. The Rsquared value indicates that the model correlates to ~98% of the variation within the training data. The MAE is also very small, which compared to the average result is ~2.7% variation. These values indicate that this model has been to extreme precision. One major concern with this high correlation is the possibility of overfitting. It will be important to assess the model on the training data.

## 4(c) - Predictive Performance on Testing Data

```{r test model on testing data}
print("Initial Neural Network on Testing Data:")
predict_funct(initial_ann, testing_data, testing_data$Y)
```

The ANN model on the testing data is also performing exceptionally well, as expected, the RMSE and MAE are higher than the training data, and the Rsquared value is lower. However, the model still correlates to ~96% of the variation within the testing data (compared to the ~98% in the training data). The MAE value, is more than twice the value of the MAE in the training data predictions (0.4220960 vs 0.1868678).

The results from both sets of predictions indicate that the ANN model, like the SVM model, performs exceptionally well. The ANN model appears to be performing better than the SVM, and certainly performs better on the current set of data.

## 4(d) - Varied Hidden Layers
```{r 1 Neuron in 1 to 10 layers}
# #multi_layer_df <- data.frame(CPUTime = numeric(0), Layers = integer(0), Neurons = integer(0), train_RMSE = numeric(0), train_R2 = numeric(0), train_MAE = numeric(0), test_RMSE = numeric(0), test_R2 = numeric(0), test_MAE = numeric(0))
# for(i in 1:10){
#   hidden_layers <- rep(1,i)
#   time_to_train <- system.time({
#     set.seed(0)
#     ann <- neuralnet(model_formula,
#                      data = training_data,
#                      hidden = hidden_layers,
#                      stepmax = 1e7
#                      )
#     })
#   train_pred <- predict_funct(ann, training_data, training_data$Y)
#   test_pred <- predict_funct(ann, testing_data, testing_data$Y)
#   new_row <- data.frame(
#     CPUTime = time_to_train[3], 
#     Layers = i,
#     Neurons = 1, 
#     train_RMSE = train_pred[1], 
#     train_R2 = train_pred[2],
#     train_MAE = train_pred[3], 
#     test_RMSE = test_pred[1],
#     test_R2 = test_pred[2],
#     test_MAE = test_pred[3])
#   multi_layer_df <- rbind(multi_layer_df, new_row)
#   
# }
# saveRDS(multi_layer_df, file = "multi_layer_df.rds")
multi_layer_df <- readRDS(file = "multi_layer_df.rds")
multi_layer_df
```
In this set of testing, it appears that an increase in the number of layers, with a constant number of neurons (in this case 1), surprisingly the model gets worse on the the training data predictions, and stays relatively stable in the testing data predictions, with a slight dip. The model training time rises, with a sharp jump at the 6 and 7 layer models, before dropping again for the 8, 9 and 10 layer models. The initial expectation was that the model would get better with a larger number of hidden layers, but this does not appear to be the case. More layers should mean that the model is better able to identify patterns within the data. The worsening results could be a result of the model overfitting the data, but that only makes sense of the predictions on the test data, it does not make sense of the training data predictions. Ultimately there does not appear to be a pattern in an increasing number of layers, with a single neuron.

## 4(e) - Varied Neurons
```{r 1 to 10 neurons in a single layer}
# #multi_neuron_df <- data.frame(CPUTime = numeric(0), Layers = integer(0), Neurons = integer(0), train_RMSE = numeric(0), train_R2 = numeric(0), train_MAE = numeric(0), test_RMSE = numeric(0), test_R2 = numeric(0), test_MAE = numeric(0))
# for(i in 1:10){
#   time_to_train <- system.time({
#     set.seed(0)
#     ann <- neuralnet(model_formula,
#                      data = training_data,
#                      hidden = c(i),
#                      stepmax = 1e7
#                      )
#     })
#   train_pred <- predict_funct(ann, training_data, training_data$Y)
#   test_pred <- predict_funct(ann, testing_data, testing_data$Y)
#   new_row <- data.frame(
#     CPUTime = time_to_train[3], 
#     Layers = 1,
#     Neurons = i, 
#     train_RMSE = train_pred[1], 
#     train_R2 = train_pred[2],
#     train_MAE = train_pred[3], 
#     test_RMSE = test_pred[1],
#     test_R2 = test_pred[2],
#     test_MAE = test_pred[3])
#   multi_neuron_df <- rbind(multi_neuron_df, new_row)
#   print(i)
# }
# saveRDS(multi_neuron_df, file = "multi_neuron_df.rds")
multi_neuron_df <- readRDS(file = "multi_neuron_df.rds")
multi_neuron_df
```
In this set of testing, a varying number of neurons has been trained on a single hidden layer. In this dataset the set of data containing 2 neurons is missing, because it would not train in the given number of steps. In this set of testing, the Rsquared value for the testing data predictions increases quite rapidly, and staying above 0.95 for most sets of testing. However the Rsquared value on the testing data jumps all around the place, usually staying below 0.4. This is an excellent indication that the high number of neurons have overfit the data.

Ultimately, the number of neurons and layers will depend on the complexity of the data it is intended to classify, and its underlying patterns. It is difficult to make assess patterns on a single set of data. However generally, a larger number of hidden layers and neurons per layer leads to better results in complex data, while in simpler datasets this can lead to overfitting.

```{r Run Until Chunk, eval=FALSE, include=FALSE}
```

# Comparison and Suggestions

## 5(a) - Cross-validated Accuracy
Cross-validated accuracy is a crucial step in machine learning, that takes the training data and tests the models against a number of subsets of the data to assess the accuracy of the model. Cross validation is done for several important reasons:
1) Cross validation provides an estimate of the goodness of fit of the model against unseen data
2) Cross validation allows the comparison of different hyperparameters to select the hyperparameters which perform the best
3) Cross validation allows for the assessment of overfitting and underfitting of the model
4) Cross validation allows for the assessment of model robustness, by testing on multiple splits of the data, this ensures that the models performance is consistent, and not due to a specific split of the training data
5) Cross validation provides a an estimate of the model's overall accuracy. This helps assess whether the model is stable or not.

Repeated cross validation was used in the hyperparameter tuning for the SVM model initially, allowing for the best cost and sigma parameters to be chosen.
Cross validation was not performed for the ANN model, which is a major limitation in trusting the model on unseen data, although its performance on the testing data is somewhat of a mitigating factor

## 5(b) - Computational Time to Train Models
```{r Computational Time}
cat("Time to train SVM (seconds): ", time_svm[3], "\n")
cat("Time to train ANN (seconds): ", time_ann[3], "\n")
cat("Time to tune SVM Parameters (seconds): ", time_svm_tuning[3], "\n")
```
One of the major considerations in choosing a machine learning model is the training time. Not only the true training time for the model, but the amount of time it takes to find the appropriate hyperparameters (tuning time). The tuning time is affected by multiple factors, including the number of observations, model type/complexity and validation measures (cross validation, bootstrapping etc.).

In this report, the time it took to train the SVM and ANN are negligible, 0.01 and 9.41 seconds respectively. This is in large part due to the small number of observations in the training dataset (<150).

The big problem, regarding computational time, is in the hyperparameter tuning. For the SVM this value is 255.75 seconds (~4 minutes) and no tuning was performed on the ANN model, but it is likely to be larger, with the same number of repeats 


1.96 hours (~117 minutes) respectively. These values can be adjusted by changing the hyperparameters to be tested, and changing the number of cross-validations. In this report proper hyperparameter tuning has not been performed on the ANN model, although it could be done in the same way as the SVM. Proper hyperparameter tuning of the ANN model, utilising cross validation, would result in a longer time to tune.

## 5(c) - Interpretability
Neither the SVM nor ANN models are particularly interpretable. We can extract the hyperparameters (C and Sigma) used in the SVM, and the defined number of Hidden Layers and Neurons used in the ANN. One of the drawbacks in this case for the SVM is the choice of kernel, a linear kernel would make the result more interpretable, as the hyperplane is more easily described, but the radial kernel used here makes this difficult, but not necessarily impossible. An ANN model is considered a black box in general, the process of back propagation, used to adjust the weights of the neurons, by assessing the predictions with the labelled data, is not easily understood.

## 6 - Suggestions for Prediction Improvement (Methodologies and Data)
### Methodology
One major improvement to the implemented methodologies would be to implement cross-validation of the ANN model, allowing better trusting of the ANN model on different data.

A second major improvement would be to increase the range of hyperparameters to be tested. This would allow better trusting of the hyperparameters chosen to train the model on.

Both of these suggested improvements will be significantly limited in implementation, by computational constraints, and number of observations.

### Data
The major improvement to the data would be by gathering more observations, and more predictors. Gathering more observations is self-explanatory, the more observations a model can be trained on, the more accurate the model can be. More predictors however, has the side effect of introducing more noise into the data. The choice of predictors is ultimately up to domain experts. The 7 predictors in this dataset may be the most appropriate predictors to use, or maybe one should be swapped out for something more relevant. Ideally, the number of predictors is minimised, but the importance of the predictors is maximized. Whether or not this is the case in this data is unknown.


# References
C, P. (2021, November 17). Missing Values Be Gone. Medium. https://towardsdatascience.com/missing-values-be-gone-a135c31f87c1

Chugh, A. (2020, December 8). MAE, MSE, RMSE, Coefficient of Determination, Adjusted R Squared — Which Metric Is Better? Medium. https://medium.com/analytics-vidhya/mae-mse-rmse-coefficient-of-determination-adjusted-r-squared-which-metric-is-better-cd0326a5697e

https://twitter.com/cassiasamp. (2023, April 21). Understanding SVM Hyperparameters. Stack Abuse. https://stackabuse.com/understanding-svm-hyperparameters/

Machine learning - when to use random forest over SVM and vice versa? (n.d.). Data Science Stack Exchange. Retrieved October 5, 2023, from https://datascience.stackexchange.com/questions/6838/when-to-use-random-forest-over-svm-and-vice-versa

More, P. (2020, January 27). An Introduction to Missing Value Imputation in Univariate Time series. Medium. https://medium.com/@poojamore282/an-introduction-to-missing-value-imputation-in-univariate-time-series-7739a34e87e3

neuralnet: Train and Test Neural Networks Using R. (2019, September 19). DataScience+. https://datascienceplus.com/neuralnet-train-and-test-neural-networks-using-r/

Shulga, D. (2018, September 27). 5 Reasons Why You Should Use Cross-Validation in Your Data Science Projects. Medium. https://towardsdatascience.com/5-reasons-why-you-should-use-cross-validation-in-your-data-science-project-8163311a1e79

# Appendices

# Leave Out Of Real Report Due To Space
## Assess Residuals

Assess Residuals There are three main assumptions for residuals in a
regression model: 1) Normal Distribution - The residuals should be
normally distributed 2) Independence - The residuals should be random 3)
Constant Variance - The residuals should have constant variance

```{r Create Residuals Function}
#resid() function in base R does not seem to work with ann model, therefore manually find residuals so that results are comparable in case of implementation difference in resid()
examine_residuals_funct <- function(model, data, target){
  x <- model
  y <- data
  z <- target
  #calculate predictions
  predictions <- predict(x, y)
  #calculate residuals
  residuals <- z - predictions
  #plots
  plot(residuals)
  hist(residuals)
  qqnorm(residuals)
  qqline(residuals)
  #statistics
  stats <- data.frame(matrix(ncol = 2, nrow = 0))
  colnames(stats) <- c("Stat", "Value")
  mean_row <- data.frame(Stat = "Mean", Value = mean(residuals))
  sd_row <- data.frame(Stat = "SD", Value = sd(residuals))
  skewness_row <- data.frame(Stat = "Skewness", Value = skewness(residuals))
  kurtosis_row <- data.frame(Stat = "Kurtosis", Value = kurtosis(residuals))
  stats <- rbind(stats, mean_row, sd_row, skewness_row, kurtosis_row)
  print(stats)
}
```

```{r Examining the Residuals of SVM model on training data}
examine_residuals_funct(svm_model, training_data, training_data$Y)
```

SVM Residuals (Training Data) 1) Normal Distribution Examination of the
histogram show the residuals appear to be fairly normally distributed,
with a couple of outliers. There might be a slight skew, although this
appears negligible. The QQ plot supports this, with the majority of the
data around the normal line, and a few obvious outliers below the -1
sample quantile and above the 1 sample quantile. The QQ plot also
supports the idea of SOMETHING TO DO WITH KURTOSIS.

Comparing the results with the calculated skewness and kurtosis values
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

2)  Independence Examining the residuals plot, there does not appear to
    be a relationship between consecutive residuals.
3)  Constant Variance Examining the residuals plot, there does appear to
    be a difference in the variance, towards the beginning of the data,
    there are a few outliers that make the variance look like this
    assumption is violated. However, these appear to be outliers, and
    the majority of the residuals do have a clear constant variance.

The 3 assumptions appear to be met.

```{r Examining the Residuals of SVM model on testing data}
examine_residuals_funct(svm_model, testing_data, testing_data$Y)
```

SVM Residuals (Testing Data) Examining the training data must be done
with extreme caution, due to the number of observations in this set
being so low.

1)  Normal Distribution Examination of the histogram show the residuals
    appear to be fairly normally distributed, with a couple of outliers.
    There does not appear to be any skew in the data. The QQ plot
    supports this, with the majority of the data around the normal line,
    and 2 obvious outliers towards the end. The QQ plot also supports
    the idea of SOMETHING TO DO WITH KURTOSIS.

Comparing the results with the calculated skewness and kurtosis values
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

2)  Independence Examining the residuals plot, there does appear to be a
    relationship between consecutive residuals, but this is likely due
    to the small number of observations, and the human brains evolution
    to detect patterns.

3)  Constant Variance Examining the residuals plot, there does appear to
    be a difference in the variance, towards the beginning of the data,
    there are 2 obvious outliers. The scale of the residuals plot should
    also be noted, most of the data lies between 0.0 and -0.4, with the
    two outliers at 0.6.

The 3 assumptions appear to be met on the testing data.

Ultimately this appears to be a good model, showing exceptional accuracy
in both the training and testing data sets, with the residuals suppoting
this.

Assess Residuals

```{r Examining the Residuals of ANN model on training data}
examine_residuals_funct(initial_ann, training_data, training_data$Y)
```

1)  Normal Distribution Examining the histogram, the data appears to be
    normally distributed, perhaps a slight skew.

2)  Independence Examining the residual plot, there does not appear to
    be a pattern in the residuals.

3)  Constant Variance The residuals appear to have a very constant
    variance

The initial Artificial Neural Network very clearly meets all three
residual assumptions.

```{r Examining the Residuals of ANN model on testing data}
examine_residuals_funct(initial_ann, testing_data, testing_data$Y)
```

As with the SVM model testing residuals, caution must be taken in
assessing the residuals due to the small number of observations. 1)
Normal Distribution Examining the histogram, the data appears to be
normally distributed, perhaps a slight skew.

2)  Independence Examining the residual plot, there does appear to be a
    pattern in the residuals.

3)  Constant Variance The residuals appear to have a very constant
    variance, given the scale and observation count.

The initial Artificial Neural Network appears to meet these three
assumptions.